Can Auxiliary Indicators Improve COVID-19 Forecasting and Hotspot Prediction?

### 2021-06-22 forecast/paper/supplement-text.tex

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness, we use the cross correlation function (CCF)
between the two time series. The $\CCF_\ell(a)$ of an indicator $X_{\ell}$ and case
rates $Y_{\ell}$ is defined as
% \[
% \CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
%   \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
%   \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
% \]
% where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
% shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
% are the sample mean and
% standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
their Pearson correlation where $X_\ell$ has been
% sequence of Pearson correlations for each selected $a$, where we
aligned
with the values of $Y_{\ell}$ that occurred $a$ days earlier.
Thus, for any $a>0$, $\CCF_{\ell}(a)>0$ indicates that
$Y_{\ell,t}$ is moving together with the values of $X_{\ell,t+a}$.
% days in the future.
In this case we say that $X_{\ell}$ is lagging
$Y_{\ell}$. For $a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is
positively correlated with $X_{\ell,t-a}$, so we say that
$X_{\ell}$ leads $Y_{\ell}$.

Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
containing Charlotte, North Carolina, from August 1, 2020 until the end of
September. These are the same signals shown in Figure 1 in the manuscript but
using finalized data. To
define ``leadingness'' we compute $\CCF_{\ell}(a)$ for each $a\in
\{-15,\ldots,15\}$ using the 56 days leading up to the target date (this is the same
amount of data used to train the forecasters: 21 days of training data, 21 days
to get the response at $a=21$, and 14 days for the longest lagged value). The
orange dashed horizontal line represents the 95\% significance threshold for
correlations based on 56 observations. Any correlations larger in magnitude than
this value are considered statistically significant under the null hypothesis of
no relationship. We define leadingness to be the sum of the significant
correlations that are leading (those above the dashed line with $a<0$) while
laggingness is the same but for $a>0$. In this case, there are no significant
correlations, so both scores will be 0: on September 28, \dv~is neither
leading nor lagging cases.

@@

In calculating the CCF and the associated leadingness and laggingness scores, we
have used the finalized data, and we look at the behavior when $t$ is the
target date of the forecast. That is we are using the same data to
evaluate predictive accuracy as to determine
leadingness and laggingness. It should be noted that the leadingness of the
indicator at the time the model is trained may also be important. Thus, we could
calculate separate leadningness and laggingness scores for the trained model and
for the evaluation data and examine their combination in some way. We do not
pursue this combination further and leave this question for future work.

>>>>

In Section 2.D of the main text, we discuss the extent to which the indicators
are leading or lagging case rates during different periods. To define the amount
of leadingness or laggingness at location $\ell$, we use the cross correlation
function (CCF) 
between the two time series. The $\CCF_\ell(a)$ of an indicator $X_{\ell}$ and case
rates $Y_{\ell}$ is defined as
% \[
% \CCF_{\ell,t}(a) = \frac{1}{n_a} \sum_{i=1}^{n_a} \left( ( X_{\ell,t,i+a} -
%   \overline{X}_{\ell,t}) / s^{X}_{\ell,t}\right) \left( ( Y_{\ell,t,i} -
%   \overline{Y}_{\ell,t}) / s^{Y}_{\ell,t}\right), 
% \]
% where $n_a$ is the number of available time points when $X_{\ell,t}$ has been
% shifted in time by $a$ days, and $\overline{X}_{\ell,t}$, $s^X_{\ell,t}$
% are the sample mean and
% standard deviation of $X_{\ell,t}$ (respectively $Y_{\ell,t}$). The result is a
their Pearson correlation where $X_\ell$ has been
% sequence of Pearson correlations for each selected $a$, where we
aligned
with the values of $Y_{\ell}$ that occurred $a$ days earlier.
Thus, for any $a>0$, $\CCF_{\ell}(a)>0$ indicates that
$Y_{\ell,t}$ is moving together with $X_{\ell,t+a}$.
% days in the future.
In this case we say that $X_{\ell}$ is lagging
$Y_{\ell}$. For $a<0$, $\CCF_{\ell}(a)>0$ means that $Y_{\ell,t}$ is
positively correlated with $X_{\ell,t-a}$, so we say that
$X_{\ell}$ leads $Y_{\ell}$.

Figure \ref{fig:ccf-dv-finalized} shows the standardized signals for the HRR
containing Charlotte, North Carolina, from August 1, 2020 until the end of
September. These are the same signals shown in Figure 1 in the manuscript but
using finalized data. To
define ``leadingness'' we compute $\CCF_{\ell}(a)$ (as implemented with the R
function {\tt ccf()}) for each $a\in
\{-15,\ldots,15\}$ using the 56 days leading up to the target date. This is the same
amount of data used to train the forecasters: 21 days of training data, 21 days
to get the response at $a=21$, and 14 days for the longest lagged value. The
orange dashed horizontal line represents the 95\% significance threshold for
correlations based on 56 observations. Any correlations larger in magnitude than
this value are considered statistically significant under the null hypothesis of
no relationship. We define leadingness to be the sum of the significant
correlations that are leading (those above the dashed line with $a<0$) while
laggingness is the same but for $a>0$. In the figure, there are three
significant correlations on the ``leading'' side (at $a = -5, -4, -3$), so
leadingness will be the sum of those values while laggingness is 0:
%correlations, so both scores will be 0:
on September 28 in Charlotte, \dv~is leading cases
leading but not lagging.

@@

In calculating the CCF and the associated leadingness and laggingness scores, we
have used the finalized data, and we look at the behavior at the 
target date of the forecast. That is we are using the same data to
evaluate predictive accuracy as to determine
leadingness and laggingness. It should be noted that the leadingness of the
indicator at the time the model is trained may also be important. Thus, we could
calculate separate leadingness and laggingness scores for the trained model and
for the evaluation data and examine their combination in some way. We do not
pursue this combination further and leave this investigation for future work.

### 2021-06-22 forecast/paper/supplement.tex

\caption{Illustration of the cross-correlation function between DV-CLI and cases. The left panel shows the standardized signals over the period from August 1 to September 28 (as of September 28, 2020). The right panel shows $\CCF(a)$ for different values of $a$ as vertical blue bars. The orange dashed lines indicate the 95\% significance threshold. By our leadingness metric, DV-CLI is neither leading nor lagging cases over this period.}\label{fig:ccf-dv-finalized}
\end{figure}

>>>>

\caption{Illustration of the cross-correlation function between DV-CLI and cases. The left panel shows the standardized signals over the period from August 1 to September 28 (as of May 15, 2021). The right panel shows $\CCF_{\ell}(a)$ for different values of $a$ as vertical blue bars. The orange dashed lines indicate the 95\% significance threshold. By our leadingness/laggingness metric, DV-CLI is leading (but notlagging) cases over this period.}\label{fig:ccf-dv-finalized}
\end{figure}
